Verifying Android Applications using Java PathFinder

### 2012-09-03 Papers/paper.tex
Updating spelling

This paper describes the development of JPF-ANDROID, an Android application verification tool. JPF-ANDROID is built on Java PathFinder, 
a Java model checking engine. JPF-ANDROID provides a simplified model of the Android framework on which an Android application can run. It then allows
the user to script input events to drive the application flow. JPF-ANDROID provides a way to detect common property violations such as deadlocks and runtime exceptions
in Android applications.
\end{abstract}

@@

Although Android applications and Java desktop applications are designed for completely different Java VMs, they are both built on an
implementation of the Java API. As a consequence, Android applications contain many of the same error as Java applications. These defects
include, but are not limited to, concurrency issues and common runtime exceptions such as NullPointer Exceptions. It follows that existing
Java testing frameworks can be adapted to verify Android applications.

% what are we going to do and why 
This paper describes an extension to Java PathFinder (JPF)~\cite{JPFDocs} that enables the automatic verification of Android applications on
the standard Java JVM.

The next section will provide an overview of JPF and Android and how JPF can be used to test Android applications. Thereafter
the development of the JPF-ANDROID tool will be discussed.

@@

% Explain what system events are how they are handled by android how they are handled by the system
System events use \texttt{Intent} objects to describe a the event that occurred. They include events to start an Activity or Service or a
battery low notification. In the Android OS, system events are sent to the Activity manager which is modelled in JPF-ANDROID by the native
\texttt{JPF\_ActivityManager} class. This class keeps a map of Intents as they are defined in the script file. When a system event is
received, the corresponding \texttt{Intent} is looked up in the map. The \texttt{Intent} is then resolved to the relevant
component and scheduled in the application's message queue.

@@

JPF-ANDROID scripts also adopted the \textbf{REPEAT} and \textbf{ANY} constructs from JPF-AWT. \textbf{REPEAT} constructs repeat a list of
events a specified number of times. \textbf{ANY} constructs take a list of events as parameters and then uses a ChoiceGenerator and JPF's
state matching and backtracking features to visit each of the execution branches. JPF stores the state of the system before the it advances
to a new state. JPF-AWT uses a JPF search listener to store and retrieve the current position of the input script in a specific state so
that the script's state is also saved.

@@

% \begin{figure}
% \centering
% \epsfig{file=newContact.eps, width=1in,}\\
% \caption{Contacts application window}
% \end{figure}
% \begin{figure}
% {\small
% {\sf
% ...\\
% \textbf{ANY} \{ \$cancelButton.onclick(), \$clearButton.onClick()  \}
% \$nameEdit.setText(``Mary'')\\
% \$createButton.onClick()\\
% ...
% }
% }
% \caption{Input script for ``New Contact'' window}
% \end{figure}
Let us take a contacts android application as an example. The application has two Activities: ListContactsActivity and the
NewContactActivity. The ListContactsActivity contains a list of the current contacts and a button to add new contacts that directs the user
to the NewContactActivity. The NewContactActivity contains two buttons: the clear button stays on the current Activity and clears its fields
and the create button starts the \texttt{ListContactsActivity}. Now let us look at figure \ref{fig:contact} containing the input sequence
for the NewContactActivity.

@@

The problem occurs in the ANY structure. When the create button is clicked the application changes to the ListContactsActivity
and the following event is not valid any more. This drastically restricts the number of sequences that can be scripted in the file. To
address this issue, we included the use of sections in the input script (see figure \ref{fig:contact2}). Each section groups the input
events of a specific Activity. Now, if the create button is pressed the \texttt{ListContactsActivity} will be started and the
\texttt{ListContactsActivity}'s events specified in its sequence will start executing again.

@@

The addition of the sections in the input script lead to infinite event sequences. In the above example, an infinite loop occurs when the
create button is pressed on the NewContactActivity window. This will stop the execution of the NewContactActivity section and restart
the ListContactsActivity section's events. When the ListContactsActivity's section is restarted, it again clicks on the createButton
restarting the NewContactActivity. To address this issue, JPF-ANDROID keeps record of its current Activity. The scripting environment was
also adapted to store the current position of each section's events for each visited Activity. If the branch returns to a previously visited
Activity, its current position in its section is looked up and instead of restarting the section, it continues from its previous position.

@@

%how they where detected
JPF is automatically configured to detect thrown exceptions and to stop execution. When JPF-ANDROID was run it firstly detected the
\texttt{ArithmenticException} due to the division by zero:
\begingroup
 \fontsize{6pt}{7pt}\selectfont
\begin{verbatim}
====================================================== results
error #1: gov.nasa.jpf.jvm.NoUncaughtExceptionsProperty 
         "java.lang.ArithmeticException: Division by zero  at ..."
====================================================== statistics
elapsed time:       00:00:01
states:             new=76, visited=0, backtracked=63, end=30
search:             maxDepth=13, constraints hit=0
choice generators:  thread=7 (signal=0, lock=3, shared ref=0), data=39
heap:               new=3455, released=2105, max live=1507, gc-cycles=74
instructions:       76796
max memory:         117MB
loaded code:        classes=145, methods=2064
======================================================
\end{verbatim}
\endgroup
After this error was fixed JPF-ANDROID detected the \texttt{NullPointerException} in the same way.

@@

This extension provides a basis on which Android applications can be tested. It can later be extended to verify functional requirements
and identify Android specific errors using JPF's listener mechanism.
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\scriptsize
\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case

>>>>

This paper describes the development of JPF-ANDROID, an Android application verification tool. JPF-ANDROID is built on Java Pathfinder, 
a Java model checking engine. JPF-ANDROID provides a simplified model of the Android framework on which an Android application can run. It then allows
the user to script input events to drive the application flow. JPF-ANDROID provides a way to detect common property violations such as deadlocks and runtime exceptions
in Android applications.
\end{abstract}

@@

Although Android applications and Java desktop applications are designed for completely different Java VMs, they are both built on an
implementation of the Java API. As a consequence, Android applications contain many of the same errors as Java applications. These defects
include, but are not limited to, concurrency issues and common runtime exceptions such as NullPointer Exceptions. It follows that existing
Java testing frameworks can be adapted to verify Android applications.

% what are we going to do and why 
This paper describes an extension to Java Pathfinder (JPF)~\cite{JPFDocs} that enables the automatic verification of Android applications on
the standard Java JVM.

The next section will provide an overview of JPF and Android and how JPF can be used to test Android applications. Thereafter
the development of the JPF-ANDROID tool will be discussed concluding with two case studies.

@@

% Explain what system events are how they are handled by android how they are handled by the system
System events use \texttt{Intent} objects to describe a event that occurred. They include events to start an Activity or Service or a
battery low notification. In the Android OS, system events are sent to the Activity manager which is modelled in JPF-ANDROID by the native
\texttt{JPF\_ActivityManager} class. This class keeps a map of Intents as they are defined in the script file. When a system event is
received, the corresponding \texttt{Intent} is looked up in the map. The \texttt{Intent} is then resolved to the relevant
component and scheduled in the application's message queue.

@@

JPF-ANDROID scripts also adopted the \textbf{REPEAT} and \textbf{ANY} constructs from JPF-AWT. \textbf{REPEAT} constructs repeat a list of
events a specified number of times. \textbf{ANY} constructs take a list of events as parameters and then uses a ChoiceGenerator and JPF's
state matching and backtracking features to visit each of the execution branches. JPF stores the state of the system before it advances
to a new state. JPF-AWT uses a JPF search listener to store and retrieve the current position of the input script in a specific state so
that the script's state is also saved.

@@

% \begin{figure}
% \centering
% \epsfig{file=newContact.eps, width=1in,}\\
% \caption{Contacts application window}
% \end{figure}
% \begin{figure}
% {\small
% {\sf
% ...\\
% \textbf{ANY} \{ \$cancelButton.onclick(), \$clearButton.onClick()  \}
% \$nameEdit.setText(``Mary'')\\
% \$createButton.onClick()\\
% ...
% }
% }
% \caption{Input script for ``New Contact'' window}
% \end{figure}
Let us take a Contacts Android application as an example. The application has two Activities: ListContactsActivity and the
NewContactActivity. The ListContactsActivity contains a list of the current contacts and a button to add a new contact that directs the
user to the NewContactActivity. The NewContactActivity contains two buttons: the clear button stays on the current Activity and clears its
fields and the create button starts the \texttt{ListContactsActivity}. Now let us look at figure \ref{fig:contact} containing the input
sequence for the NewContactActivity.

@@

The problem occurs in the ANY structure. When the create button is clicked the application changes to the ListContactsActivity
and the following event is not valid any more. This drastically restricts the number of sequences that can be scripted in the file. To
address this issue, we included the use of sections in the input script (see figure \ref{fig:contact2}). Each section groups the input
events of a specific Activity. Now, if the create button is pressed the \texttt{ListContactsActivity} will be started and the
\texttt{ListContactsActivity}'s events specified in its section will start executing again.

@@

The addition of the sections in the input script lead to infinite event sequences. In the above example, an infinite loop occurs when the
create button is pressed on the NewContactActivity window. This will stop the execution of the NewContactActivity section and restart
the ListContactsActivity section's events. When the ListContactsActivity's section is restarted, it again clicks on the createButton
restarting the NewContactActivity. To address this issue, JPF-ANDROID keeps record of its current Activity. The scripting environment was
also adapted to store the current position in each visited Activity's script section. If a branch returns to a previously visited
Activity, its current position in its section is looked up and instead of restarting the section, it continues from its previous position.

@@

%how they where detected
JPF is automatically configured to detect thrown exceptions and to stop execution. When JPF-ANDROID was run on the application it firstly
detected the \texttt{ArithmenticException} due to the division by zero:
\begingroup
 \fontsize{6pt}{7pt}\selectfont
\begin{verbatim}
====================================================== results
error #1: gov.nasa.jpf.jvm.NoUncaughtExceptionsProperty 
         "java.lang.ArithmeticException: Division by zero  at ..."
====================================================== statistics
elapsed time:       00:00:01
states:             new=76, visited=0, backtracked=63, end=30
search:             maxDepth=13, constraints hit=0
choice generators:  thread=7 (signal=0, lock=3, shared ref=0), data=39
heap:               new=3455, released=2105, max live=1507, gc-cycles=74
instructions:       76796
max memory:         117MB
loaded code:        classes=145, methods=2064
======================================================
\end{verbatim}
\endgroup
After this error was fixed JPF-ANDROID detected the \texttt{NullPointerException} in the same way.

@@

This extension provides a basis on which Android applications can be tested. It can later be extended to verify functional requirements
and identify Android specific errors using JPF's listener mechanism.
\vspace*{5pt}
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\scriptsize
\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
